\documentclass{article}
\usepackage[UTF8]{ctex}
\usepackage[margin=0.7in]{geometry}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{graphicx}


\begin{document}

\begin{titlepage}
    \centering
    \vspace*{2cm}
    
    {\Huge\bfseries Building LLM Reasoners\par}
    \vspace{0.5cm}
    {\Large Assignment 2\par}
    \vspace{2cm}
    
    {\Large\bfseries Building a Transformer LM\par}
    \vspace{3cm}
    
    {\Large Siva Satvik\par}
    \vspace{0.5cm}
    {\large sm12779\par}
    \vspace{2cm}
    
    \vfill
\end{titlepage}

\begin{enumerate}[start=1,label=\arabic*.]
    \item \textbf{Profiling and Benchmarking}
    \begin{enumerate}
        \item[1.1.3] \textbf{\texttt{benchmarking\_script}}
            \begin{enumerate}[label=\alph*.]
                \item The script is present at student/benchmark.py
                \item The following table shows the timings for the forward and forward+backward passes for each of the model sizes. The mean and standard deviation are computed across 10 runs for each model size and pass type.
                    \begin{table}[h]
                        \centering
                            \begin{tabular}{|r|r|r|r|}
                                \hline
                                model\_size & pass & mean\_step\_ms & std\_step\_ms \\
                                \hline
                                small & forward & 34.7421 & 0.2440 \\
                                \hline
                                small & forward+backward & 84.1578 & 0.4178 \\
                                \hline
                                medium & forward & 71.0755 & 3.2061  \\
                                \hline
                                medium & forward+backward & 215.5571 & 0.3645  \\
                                \hline
                                large & forward & 141.2933 & 0.1184 \\
                                \hline
                                large & forward+backward & 436.3411 & 0.3570  \\
                                \hline
                                xl & forward & 280.6156 & 0.0603 \\
                                \hline
                                xl & forward+backward & 888.9120 & 0.1919  \\
                                \hline
                                2.7B & forward & 417.7063 & 0.1057 \\
                                \hline
                                2.7B & forward+backward & 1320.3845 & 1.0714  \\
                                \hline
                            \end{tabular}
                    \end{table}
                    Backward pass is taking significantly more time than the forward pass, which is expected since it involves computing gradients for all parameters. As the model size increases, both forward and backward pass times increase, with the backward pass showing a more significant increase due to the larger number of parameters and computations involved. Generally, variance in timings is low (medium forward pass has higher variance than others), which suggests that the measurements are consistent across runs.
                \item When running the measurement steps without the warmup steps, the timings for the first couple of iterations are significantly higher than the subsequent iterations. This is obvious when running the measurement steps with 2 warmup steps, where the timings are almost same as runs with 5 warmup steps as in the previous experiment. This is because the first few iterations involve additional overhead from things like memory allocation, kernel compilation, and other setup tasks that are not representative of the steady-state performance of the model. By including warmup steps, we allow the system to reach a more stable state before we start measuring the performance, which results in more accurate and consistent timing measurements.
            \end{enumerate}
        \item[1.1.4] \textbf{\texttt{nsys\_profile}}
            \begin{enumerate}[label=\alph*.]
                \item The times for forward pass here are very similar to the times obtained in the previous section for the forward pass. This is expected since both measurements are for the same operation (forward pass) and should be consistent (The context length is the same in the comparision).
                \item The CUDA kernel that takes the most time in the forward pass is the ampere\_sgemm\_128x64\_tn kernel, which is a matrix multiplication kernel used in the transformer block. This kernel is responsible for Q, K, V and output matrix multiplications in attention layer and feed forward layers weights computations. Per transformer block, there are 7 calls to this kernel and for one forward pass, it will be number of blocks * 7 calls to this kernel. In both forward and backward pass, there's another kuda kernel that takes almost similar time as the ampere\_sgemm\_128x64\_tn kernel, which is the ampere\_sgemm\_128x128\_nt kernel. This kernel is also a matrix multiplication kernel used in the transformer block, and it is responsible for the matrix multiplications in the feed forward layers and attention output computations. The presence of these two kernels as the most time-consuming ones in both forward and backward passes indicates that matrix multiplication operations are the primary bottleneck in the performance of the transformer model.
                \item
                \item
                \item
            \end{enumerate}
        \item[1.1.5] \textbf{Mixed Precision}
            \begin{itemize}
                \item \textbf{\texttt{mixed\_precision\_accumulation\_test}}\\
                The output of the given code is as follows:
                    \begin{verbatim}
                        tensor(10.0001)
                        tensor(9.9531, dtype=torch.float16)
                        tensor(10.0021)
                        tensor(10.0021)
                    \end{verbatim}
                \textbf{Case 1: Pure Float32 Accumulation}
                In this scenario, both the accumulator and the increment use 32-bit single precision, which provides approximately 7 decimal digits of resolution. This is sufficient to represent the value 0.01 accurately enough that the final sum reaches 10.0001, showing nearly perfect alignment with the theoretical result.\\
                \textbf{Case 2: Pure Float16 Accumulation}
                Using 16-bit half precision significantly limits the available bits for the fraction, resulting in only about 3 decimal digits of precision. As the sum grows, the "gap" between representable numbers becomes larger than the 0.01 increment, causing the addition to lose precision or stall entirely, which leads to the inaccurate result of 9.9531.\\
                \textbf{Case 3: In-place Mixed Precision Addition}
                This case uses a Float32 sum but attempts to add a Float16 tensor directly to it, which triggers implicit type promotion. While the sum is more stable than Case 2, the increment itself was already "quantized" to a less accurate Float16 version of 0.01, resulting in a final sum (10.0021) that carries that initial approximation error through 1,000 steps.\\
                \textbf{Case 4: Explicit Float32 Casting}
                This implementation explicitly casts the Float16 increment to Float32 before performing the addition to ensure the operation occurs in a high-precision space. The result is identical to Case 3 (10.0021) because the bottleneck is the original creation of the value in Float16; once the precision is lost during the initial tensor creation, casting it to Float32 cannot recover the original ``perfect'' decimal value.\\
                \item \textbf{benchmarking\_mixed\_precision}
                    \begin{enumerate}[label=\alph*.]
                        \item The data types for each of the components is given below:
                            \begin{itemize}
                                \item the model parameters within the autocast context: \textbf{float32}
                                \item ToyModel.fc1: \textbf{float16}
                                \item ToyModel.ln: \textbf{float32} (Layer norm is kept in higher precision)
                                \item the models's predicted logits: \textbf{float16}
                                \item the loss: \textbf{float32} (loss is kept in higher precision for stability)
                                \item the model's gradients: \textbf{float32} (gradients are kept in higher precision for stability)
                            \end{itemize}
                    \end{enumerate}
                \item Layer normalization is sensitive in its reduction-heavy steps: computing mean/variance, subtracting the mean, squaring, and reciprocal square-root can lose stability in FP16 due to limited dynamic range and precision. BF16 has much larger exponent range (similar to FP32), so overflow/underflow risk is lower than FP16, but mantissa precision is still limited. In practice, many systems still keep LayerNorm in FP32 for robustness, though BF16 usually requires fewer special-case protections than FP16.
            \end{itemize}
        \item 
    \end{enumerate} 
\end{enumerate}

\end{document}